---
title: "Assignment 2 - Survey Data Analysis"
author: 
 - "Haider, Tobias Abraham"
 - "Selenge, Tuvshin"
 - "Crespí, Ahmad Antonio"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    toc: TRUE
---

# Preparation

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Imports

```{r imports, warning=FALSE, message=FALSE}
library(knitr)
library(readr)
library(stringr)
library(tidyr)
library(pillar)
library(dplyr)
library(ggplot2)
```

## Global variables

```{r gloabl_variables}
data_path <- "./data/survey_data_clean.csv"
```

## Data loading

```{r data_loading, message=FALSE}
survey_data <- read_csv(data_path)
```

## Data Cleaning

### Manual Data Cleaning
Most of the data cleaning was done by hand. The reason for this is the variety
of inconsistencies in the collected textual data. However, the manual cleaning
was done to not change any meaning of the data.

Changes data were made by hand:

 - Align nationalities to be the country name (Austrian -> Austria, Pakistani -> Pakistan)
 - Extract desired degree and university from the study program column. If any of these were not specified, it was left blank
 - Remove abbreviations in country names and study programs (A -> Austria, AUT -> Austria)
 - Ensure consistent capitalisation (Data science -> Data Science)
 - Remove invalid responses (age of 4 years was set to empty)
 - Sleep hours that contained dashed were set to a floating point number (6-7 -> 6.5)

### Data Frame Structure

To be able to use the table with all common libraries, we ensured to represent the values as their 
natural data type. All variables are factors, except for the daily hours of sleep.
 
```{r data_cleaning_names}
names(survey_data) <- c("gender", "age", "study_program", "desired_degree", "university_name", "daily_sleep", "contentness", "primary_nationality")
```

```{r data_cleaning_types}
survey_data <- survey_data %>%
  mutate(
    gender = as.factor(gender),
    study_program = as.factor(study_program),
    desired_degree = as.factor(desired_degree),
    university_name = as.factor(university_name),
    contentness = factor(contentness, levels = c("Not at all", "Slightly content", "Quite content", "Very content"), ordered = TRUE),
    primary_nationality = as.factor(primary_nationality)
  )
```

### Grouping of Study Programs

While plotting, it was notable that some verbose descriptions of study programs.
Luckily, most responses can be grouped into more general categories of study programs.
This leads to some data loss but improves the expressiveness of all analysis done afterwards.
Above all, plots are less noisy.

```{r data_cleaning_study_program_grouping}
survey_data <- survey_data %>%
  mutate(
    study_program = case_when(
      str_detect(study_program, regex("Management", ignore_case = TRUE)) ~ "Management",
      str_detect(study_program, regex("Mathematics", ignore_case = TRUE)) ~ "Mathematics",
      study_program == "Außerordentliche Studentin" ~ NA,
      TRUE ~ study_program
    )
  )
```

# Data Exploration

Before performing the statistical analysis, we can take a look at the responses
and visualize some of the values collected.

```{r data_exploration_glimpse}
glimpse(survey_data)
```

## Gender Ratio

```{r data_exploration_gender}
ggplot(survey_data, aes(x = gender, fill = gender)) +
  geom_bar() +
  theme_minimal() + 
  guides(fill="none") + 
  labs(title = "Gender Ratio", x = "Gender", y = "Number of Students")
```
## Age distribution plot

```{r data_exploration_age, warning=FALSE}
ggplot(survey_data, aes(x = age)) +
  geom_bar() +
  geom_vline(aes(xintercept = median(age, na.rm = TRUE)), linetype = "dotted", size = 1) +
  theme_minimal() + 
  labs(title = "Age Distribution", x = "Age", y = "Number of Students")
```

The median of the student age is 26, while almost all studens are between the 
age of 20 and 30 years.

## Types of Study Programs

```{r data_exploration_study_programs}
ggplot(survey_data, aes(x = reorder(study_program, study_program, function(x) - length(x)))) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.3, size = 3.5) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Types of Study Programs", x = "Study Program", y = "Number of Students")
```

This plot clearly shows that Data Science is the study program with most students
attending the course.

## Different Nationalities

```{r data_exploration_primary_nationality}
ggplot(survey_data, aes(x = reorder(primary_nationality, primary_nationality, function(x) - length(x)))) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.3, size = 3.5) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Nationalities of Students", x = "Nationality", y = "Number of Students")
```

When ordering the number of students per nationality, it can be seen that Austria
is not the top country. Many students from Pakistan attend this course. The majority of
students have nationalities from Central and Eastern Europe.

## Measure of Contentness

```{r data_exploration_contentness}
ggplot(survey_data, aes(x = contentness, fill = contentness)) +
  geom_bar() +
  theme_minimal() + 
  guides(fill="none") + 
  labs(title = "Contentness Responses Across Students", x = "Contentness", y = "Number of Students")
```

When looking at the contentness isolated from all other variables, it becomes apparent
that studens are overall content. Around two thirds are quite or very content and 
one third is only slightly or not content at all.

## Distribution of Sleep Hours per Night

```{r data_exploration_daily_sleep}
ggplot(survey_data, aes(x = daily_sleep)) +
  geom_bar() +
  geom_vline(aes(xintercept = median(daily_sleep, na.rm = TRUE)), linetype = "dotted", size = 1) +
  theme_minimal() + 
  labs(title = "Distribution of Hours of Sleep per Night", x = "Hours Sleep", y = "Number of Students")
```

Students sleep around 7 hours per night. There is some variation present in the
data. 6, 7, and 8 are the most common response.

# Visualization of Hypotheses

Next, we represented our hypotheses in the form of plots to see whether there
are some obvious patterns.

## Visualization of Relation Between Nationality and Contentness

```{r visualization_nationality_contentness}
ggplot(survey_data, aes(x = reorder(primary_nationality, primary_nationality, function(x) - length(x)), fill = contentness)) +
  geom_bar(position = "fill") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Contentness Across Nations", x = "Nationality", y = "Percentage of Students")
```

Students from Germany, Finland, Austria and Pakistan seem to be the happiest
amongst all. However, we have to note that most contries do not have sufficient
responses to actually compare them. 

```{r visualization_austrian_non_austrian_contentness}
survey_data %>%
  mutate(
    primary_nationality = ifelse(primary_nationality == "Austria", "Austria", "Other")
  ) %>%
  ggplot(aes(x = reorder(primary_nationality, primary_nationality, function(x) - length(x)), fill = contentness)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Contentness of National and International Students", x = "Nationality", y = "Percentage of Students")
```
This plot compares the contentness of Austrians specifically to international 
students. According to the data, Austrians feel slightly more content than
international students. It is notable, that unhappy students are more likely to
decline to disclose their nationality. Because of this, there is some bias in
the collected data.

## Visualization of Relation Between Sleep and Contentness

```{r visualization_sleep_contentness}
ggplot(survey_data, aes(x = daily_sleep, fill = contentness)) +
  geom_bar() +
  theme_minimal() + 
  labs(title = "Sleep and Relation to Contentness", x = "Hours Sleep", y = "Number of Students")
```

There is no clear pattern regarding the hours of sleep and contentness. We see
an increase between the ratio of quite to slightly content when the hours of
sleep is higher. The change could also be completely random.

## Visualization of Relation Between Nationality and Study Program

```{r visualization_austrian_non_austrian_study_program}
survey_data %>%
  mutate(
    primary_nationality = ifelse(primary_nationality == "Austria", "Austria", "Other")
  ) %>%
  ggplot(aes(x = reorder(primary_nationality, primary_nationality, function(x) - length(x)), fill = study_program)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Preferred Study Programs of National and International Students", x = "Study Program", y = "Percentage of Students")
```

Interestingly, there is not a single computer science student with a nationality
other than Austrian. Overall, the plot is not displaying anything surprising.
Data Science is the biggest group of students attending this cours
across all nationalities.

# Descriptive Statistics

## Hours of Sleep

```{r descriptive_statistics_daily_sleep}
summary(survey_data$daily_sleep)
```

# Analytical Statistics

## Test for Relation Between Nationality and Contentness

## Test for Relation Between Sleep and Contentness

```{r test_relation_sleep_contentness}
anova_result <- aov(daily_sleep ~ contentness, data = survey_data)
summary(anova_result)
```

## Test for Relation Between Nationality and Study Program

# Conclusion
